---
title: "Data Preparation Report - Knotweed's Tarping Survey (Dusz et al., 2020)"
author: "François-Marie Martin"
date: "11/27/2020"
output:
   html_document:
     number_sections: yes
     toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# - Reproductibility
## - System information for my session  

```{r session information}
sessionInfo()
```
  
  



# - Data visualisation and exploration  
## - Data summary

First, let's take a quick look at our dataset. For the meaning of the variables, please refer to the attached **documentation**.

```{r data import and summary, message=FALSE, comment=NA}
library(jk.dusz.tarping)
zzz_mydata_cleaned <- jk.dusz.tarping::clean_my_data()

summary(zzz_mydata_cleaned)
```





## - Data exploration following Zuur *et al.* (2010)

In order to build valid and meaningful models to explain or predict the success/failure of tarping operations for the control of Japanese knotweed *s.l.*, we need first to learn more about our data. Therefore, we thoroughly examined our variables (e.g. outliers, distribution, interactions, colinearity, variance homogeneity, observation independence) following the protocol presented by Zuur *et al.* (2010) for each of our **7 response variables**: *eff_eradication*, *eff_expansion*, *eff_vigour*, *latest_reg_edges*, *reg_overlaps*, *latest_condition*, and *pb_fixation*.


### - For the *eradication efficiency* response variable

```{r erad import, include=FALSE}
erad <- jk.dusz.tarping::model_datasets(response.var = "efficiency")
erad %>% dplyr::select(-eff_expansion, -eff_vigour, -goals) %>%
  dplyr::filter(eff_eradication != "NA") -> erad
```


#### - Looking for outliers

We will look for potential outliers in the data using **boxplots** and **Cleveland dotplots**.

    


